\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-03-18 paper.qmd
Added PBMC author's model error

trimesh_removed_pbmc_umap <- trimesh_removed_pbmc_umap +
  geom_point(data = umap_pbmc_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.3) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) #+
  #annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

# df_bin_train <- df_bin_pbmc
# names(df_bin_train)[-1] <- paste0("avg_", names(df_bin_train)[-1])
# 
# df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), umap_data_with_hb_id_pbmc)
# 
# error_df <- df_all_pbmc |>
#   dplyr::left_join(df_bin_train, by = c("hb_id" = "hb_id")) ## Map high-D averaged/weighted mean coordinates
# 
# for (i in 1:(NCOL(df_bin_train) - 1)) {
#   
#   error_df[ , paste0("abs_residual_", "PC_", i)] <- abs(error_df[ , paste0("PC_", i)] - error_df[ , paste0("avg_", "PC_", i)])
#   
# }
# 
# error_df <- error_df |>
#   dplyr::mutate(total = rowSums(dplyr::pick(tidyselect::starts_with(paste0("abs_residual_", "PC_")))))
# 
# 
# error_df <- error_df |>
#   mutate(type = if_else(total == 0, "no error",
#                         if_else(total <= 15, "error 1-15",
#                                 if_else(total <= 20, "error 15-20",
#                                         if_else(total <= 25, "error 20-25",
#                                                 if_else(total <= 30, "error 25-30",
#                                                         if_else(total <= 35, "error 30-35",
#                                                                 if_else(total <= 40, "error 35-40", "error greter than 40"))))))))
# 
# plot_list2_pbmc_error <- error_df |>
#   mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
#     ggplot(aes(x = UMAP1,
#                y = UMAP2, color = type))+
#     geom_point(alpha=0.5, size = 0.1) +
#     coord_equal() +
#     theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
#   theme_linedraw() +
#     theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
#               axis.title.x = element_blank(), axis.title.y = element_blank(),
#               axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#               axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#               panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
#         legend.title = element_text(size=5), #change legend title font size
#         legend.text = element_text(size=4),
#          legend.key.height = unit(0.25, 'cm'),
#          legend.key.width = unit(0.25, 'cm')) +
#   scale_color_manual(values=c("#fee0d2", "#fcbba1",
#                                          "#fc9272", "#fb6a4a", "#ef3b2c",
#                                          "#cb181d", "#a50f15", "#99000d")) 
```

```{r}
#| echo: false
#| fig-cap: "Model in the 2D space overlaid on UMAP data."
#| label: fig-nldervisPBMCUMAP
#| fig-pos: H

trimesh_removed_pbmc_umap 
```

>>>>

@@

plot_list2_pbmc_error <- error_df |>
  mutate(type = factor(type , levels = c("no error", "error 1-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error 35-40", "error greter than 40"))) |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d")) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

```{r}
#| echo: false
#| fig-cap: "(a) Model in the 2D space overlaid on UMAP data, and (b) Error occurred by the high-D model for high-D data and the error mapped into 2D NLDR points."
#| label: fig-nldervisPBMCUMAP
#| fig-pos: H

trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
plot_layout(guides='collect', ncol=2) &
  theme(legend.position='bottom') 
```

### 2024-03-18 paper.tex
Added PBMC author's model error

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-nldervisPBMCUMAP}Model in the 2D space overlaid on
UMAP data.}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-nldervisPBMCUMAP}(a) Model in the 2D space overlaid
on UMAP data, and (b) Error occurred by the high-D model for high-D data
and the error mapped into 2D NLDR points.}
